### 2016-01-30 Documents/paper.Rmd
initial commit

---
title: "Issues in developing mulitvariable models for treatment selection"
author: "Michael C Sachs and Lisa M McShane"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
    html_document:
      pandoc_args: ["--bibliography", "references.bib"]
---

# Introduction

## Predictive biomarkers

In what ways can it change clinical practice?

@janes2011measuring, @janes2014approach,  @mcshane2013development, @polley2013statistical

## Overview of biomarker signatures

- Do I include all features or a subset?
    + Which subset?
    + How do I combine the features?
- Transformations?
- Weight and combine with regression coefficients?
- What coefficients?
- Thresholds/Cutoffs before or after combining?

References on developing signatures @simon2003pitfalls, @hastie2009elements, @moons2012riskI, @moons2012riskII

## Data analysis example

@zhu2010prognostic

# Issues

## Estimating Prediction Error

## Estimating Clinical Benefit

## Developing and validating a predictive biomarker for an effective treatment

Identify the non-responders and demonstrate that they do not benefit (hard)

## Developing and validating a predictive biomarker for an ineffective treatment

Identify the responders and demonstrate that they do benefit (no quite as hard)

# Conclusion