### 2016-07-05 Code/new sims/supplement.Rmd
new simulations

---
title: "Issues in developing multivariable molecular signatures for guiding clinical care decisions: supplemental simulation results"
author: "Michael C Sachs and Lisa M McShane"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
    pdf_document:
      citation_package: natbib
      keep_tex: yes
      fig_caption: yes
      pandoc_args: ["--bibliography", "paper.bib"]
header-includes:
    - \usepackage{setspace}
fontsize: 11pt
geometry: margin=1in
---

# Additional Simulation Results

In the tables and figures below, $N$ is to the number of samples, and $p$ is the number of variables per sample. All results are from 1000 simulation replicates.

```{r setup, include = FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(survival)
library(ggkm)
library(tidyr)
library(stargazer)

opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE)
knit_hooks$set(plot = function(x, options) {
  if ('tikz' %in% options$dev && !options$external) {
    hook_plot_tex(x, options)
  } else hook_plot_md(x, options)
})

source("newsims.R")

```

```{r cycle, echo = FALSE, results = "asis"}
files <- grep("*.txt", list.files(), value = TRUE)
results <- lapply(files, parse_sims)

for(i in results){

  cat(i$table)

}
```

```{r plotsauc, echo = FALSE, results = "asis"}
for(i in results){

  print(i$plot.auc)
  cat("\n\\clearpage\n\n")

}
```

```{r plots2, echo = FALSE, results = "asis"}
for(i in results){

  print(i$plot.or)
  cat("\n\\clearpage\n\n")

}
```

```{r overall, fig.height = 8.5}
alldat <- NULL
for(i in results){

  alldat <- rbind(alldat, i$datalong)

}

 labs <- c("Resubstitution", "Partial CV",
            "Partial Holdout", "Partial Resubstitution",
            "Pre-validation", "Leave 10 out CV", "Leave 100 out CV",
            "30% Holdout", "50% Holdout", "Bootstrap")

  names(labs) <- levels(alldat$scen)

ggplot(subset(alldat, stat == "AUC"), aes(y = value, x = scen)) + geom_violin(fill = "grey60") +
    theme_bw(base_size = 13, base_family = "serif") + coord_flip() +
    scale_x_discrete("Estimation Approach", labels = labs) +
    ylab("Area Under the ROC Curve") + facet_grid(p ~ N, labeller = "label_both")
```

```{r, fig.height = 8.5}
alldat$value[alldat$stat == "OR"] <- exp(alldat$value[alldat$stat == "OR"])

ggplot(subset(alldat, stat == "OR"), aes(y = value, x = scen)) + geom_violin(fill = "grey60") +
    theme_bw(base_size = 13, base_family = "serif") + coord_flip() +
    scale_x_discrete("Estimation Approach", labels = labs) + scale_y_log10(breaks = c(.5, 2, 5, 20, 80)) +
    ylab("Odds Ratio (log scale)") + facet_grid(p ~ N, labeller = "label_both")

```